---
title: "Reproducible Research Project"
author: "Sangwon Han"
date: "2018³â 11¿ù 29ÀÏ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. read and process data
##1.1. Unzip and read the data
```{r}
unzip("repdata%2Fdata%2Factivity.zip")
activity <- read.csv("activity.csv")
str(activity)
```

##1.2. Transform the time data into data format
```{r}
library(lubridate)
activity$date<-ymd(activity$date)
```


#2. Histrogram of the total number of steps taken each day
##2.1. Calculate the total number of steps taken per day
```{r}
library(dplyr)
activity_date <- aggregate(steps ~ date, activity, sum)
summary(activity_date)
```

##2.2. Histrogram of the total number of steps taken each day
```{r}
hist(activity_date$steps, main = "Total number of steps taken each day", xlab = "Steps per day", col = "blue", breaks = seq(0,25000, by = 2500), ylim = c(0,20))
```


#3. Mean and median number of steps taken each day
```{r}
activity_mean <- aggregate(steps ~ date, activity, mean, na.rm = T)
activity_mean
activity_median <- aggregate(steps ~ date, activity, median, na.rm = T)
activity_median
```


#4. Time series plot of the average number of steps taken
##4.1. Make a time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
activity_interval <- aggregate(steps ~ interval, activity, mean, na.rm = TRUE)
with(activity_interval, plot(interval, steps, type = "l", col = "blue", xlab = "Interval", ylab = "Average number of steps", main = "Average number of steps per 5-min intervals"))

```


#5 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
activity_interval[which.max(activity_interval$steps),]$interval
```


#6. Code to describe and show a strategy for imputing missing data
##6.1. Calculate the total number of missing values in the dataset 
```{r}
sum(is.na(activity$steps))
```
##6.2. Devise a strategy for filling in all of the missing values in the dataset. (I filled the missing values with the average number of step per intervals)
```{r}
imputing <- activity_interval$steps[match(activity$interval, activity_interval$interval)]
str(imputing)
```
##6.3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputing, no = activity$steps))
```
##6.4.1. Make a histogram of the total number of steps taken each day 
```{r}
activity_imputed_date <- aggregate(steps ~ date, activity_imputed, sum)
hist(activity_imputed_date$steps, main = "Total number of steps taken each day after imputing missing values", xlab = "Steps per day", col = "blue", breaks = seq(0,25000, by = 2500), ylim = c(0,30))

```
##6.4.2 Calculate and report the mean and median total number of steps taken per day. 
```{r}
mean(activity_imputed_date$steps)
median(activity_imputed_date$steps)
```


#7. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
##7.1. Create a new factor variable in the dataset with two levels ??? ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.
```{r}
Sys.setlocale(category = "LC_TIME", "English")
table(weekdays(activity_imputed$date))
activity_imputed$weekday <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) == "Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
table(activity_imputed$weekday)
```
##7.2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
activity_imputed_weekdays <- aggregate(steps ~ interval + weekday, activity_imputed, mean)
library(ggplot2)
ggplot(activity_imputed_weekdays, aes(x = interval , y = steps, color = weekday)) +
        geom_line() +
        labs(title = "Average number of steps across all weekday or weekend", x = "Interval", y = "Average number of steps") +
        facet_grid(weekday~., scale = "free")
```