---
title: "Reproducible Research Project 2"
author: "Sangwon Han"
date: "2018<eb>ÖÑ 12<ec>õî 27<ec>ùº"
output:
  pdf_document: default
  html_document: default
---

#Synopsis
The basic goal of this project is to explore the NOAA Storm Database and answer some basic questions about severe weather events. 
The data analysis is for addressing the following questions:

1. Across the United States, which types of events are most harmful with respect to population health?

2. Across the United States, which types of events have the greatest economic consequences?

#Read and preprocess data
##Set working directory and read the data
```{r, cache=TRUE, results=FALSE}
getwd()
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, "StormData.csv.bz2")
df <- read.csv("StormData.csv.bz2", header=TRUE, sep=",")
```

##Explore and clean data
```{r, cache=TRUE}
str(df)
```

```{r, cache=TRUE}
head(df)
```

```{r, cache=TRUE}
tail(df)
```
There must be an error in the dataset, so find the location of the source of error in "STATE" colunm (having 2 character, normally)
```{r, cache=TRUE}
sum(nchar(as.character(df$STATE)[1:570290])==2)
```
Something wrong in 547364th row
```{r, cache=TRUE}
View(df[547363:547365,])
```
The problem exist in [END_RANGE] in 547364th row. 
```{r, cache=TRUE}
df$END_RANGE[547364]
```
Load data using [readLines] function, again. (The quote mark is &#92;")
```{r, cache=TRUE}
df1 <- readLines(con="StormData.csv.bz2")
df1[1:2]
```
Find the line having problem
```{r, cache=TRUE}
grep("547303.00", df1)
```

```{r, cache=TRUE}
df1[884718]
```
The problem is [&#92;"] in the sentences. Therefore, let's remove it
```{r, cache=TRUE}
tmp1 = gsub("\"\" ","", df1)
writeLines(tmp1, "StormData1.txt")
df_edited <- read.table("StormData1.txt", sep=",", quote = "\"", header = T)
dim(df_edited)
```


#Process data 

##Extract variables for health and economic data
```{r, cache=TRUE}
health <- c("EVTYPE", "FATALITIES", "INJURIES")
healthData <- df_edited[,health]
economic <- c("EVTYPE", "CROPDMG", "CROPDMGEXP", "PROPDMG", "PROPDMGEXP")
economicData <- df_edited[,economic]
```

##Process health data
```{r, cache=TRUE}
str(healthData)
```

```{r, cache=TRUE}
health_event <- aggregate(cbind(FATALITIES, INJURIES) ~ EVTYPE, data=healthData, FUN=sum, na.rm = TRUE)
health_event$TOTAL <- health_event$FATALITIES + health_event$INJURIES
```
arrange data in descending order, extract top 10 events, and use [melt] function for grouped bar chart 
```{r, cache=TRUE}
library(dplyr)
health_event <- arrange(health_event, desc(TOTAL))
top10_health <- health_event[1:10, ]
top10_health
```

```{r, cache=TRUE}
library(reshape2)
top10_health_melt <- melt(top10_health, id.vars = "EVTYPE")
```




#Draw bar chart for health impact
```{r, cache=TRUE}
library(ggplot2)
fig_health <- ggplot(top10_health_melt, aes(x = reorder(EVTYPE, -value), y = value)) +
                geom_bar(stat = "identity", aes(fill = variable), position = "dodge") +
                xlab("Events") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                ylab("Total counts") + 
                ggtitle("Top 10 weather events for health effect")
fig_health
```

[TORNADO] is the most harmful with respect to population health!!




## Process economic data

```{r, cache=TRUE}
table(economicData$CROPDMGEXP)
```

```{r, cache=TRUE}
table(economicData$PROPDMGEXP)
```

```{r, cache=TRUE}
str(economicData$CROPDMGEXP)
```

change character into numeric values in [CROPDMGEXP] and [PROPDMGEXP] columns. At first, I made a function 
```{r, cache=TRUE}
trans <- function(x){
        if(x=="?"|x=="-"|x=="+"|x==""){
                x <- 1
        }else if(x=="h"|x=="H"){
                x <- 10^2
        }else if(x=="k"|x=="K"){
                x <- 10^3
        }else if(x=="m"|x=="M"){
                x <- 10^6
        }else if(x=="b"|x=="B"){
                x <- 10^9
        }else{
                x <- 10^(as.numeric(x))
        }
}
```
Transform the values
```{r, cache=TRUE}
economicData$CROPDMGEXP_trans <- sapply(as.character(economicData$CROPDMGEXP), trans)
table(economicData$CROPDMGEXP_trans)
```

```{r, cache=TRUE}
economicData$PROPDMGEXP_trans <- sapply(as.character(economicData$PROPDMGEXP), trans)
table(economicData$PROPDMGEXP_trans)
```

calculate the sum of the values
```{r, cache=TRUE}
economicData$CROP_total <- economicData$CROPDMGEXP_trans * economicData$CROPDMG
economicData$PROP_total <- economicData$PROPDMGEXP_trans * economicData$PROPDMG
```

I calculate the sum of each value according to the type of events.
Then, arrange data in descending order, extract top 10 events, and use [melt] function for grouped bar chart
```{r, cache=TRUE}
economic_event <- aggregate(cbind(CROP_total, PROP_total) ~ EVTYPE, data = economicData, FUN=sum, na.rm = TRUE)
economic_event$SUM <- economic_event$CROP_total + economic_event$PROP_total
economic_event <- arrange(economic_event, desc(SUM))
top10_economic <- economic_event[1:10, ]
top10_economic
```

```{r, cache=TRUE}
colnames(top10_economic) <- c("EVTYPE", "CROP", "PROPERTY", "TOTAL")
top10_economic_melt <- melt(top10_economic, id.vars = "EVTYPE")
```




#Draw bar chart for Economic impact
```{r, cache=TRUE}
fig_economic <- ggplot(top10_economic_melt, aes(x = reorder(EVTYPE, -value), y = value)) +
        geom_bar(stat = "identity", aes(fill = variable), position = "dodge") +
        xlab("Events") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ylab("Total counts") + 
        ggtitle("Top 10 weather events for economic effect")
fig_economic
```


Therefore, [FLOOD] is the event having the greatest economic consequence!!