---
title: "Statistical Inference - course project part 1"
author: "Sangwon Han"
date: '2018-12-18'
output:
  word_document: default
  pdf_document: default
graphics: yes
html_document: default
fig_width: 5
fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=65),tidy=TRUE)
```


# Overview

In this project, i will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

The followings are the instruction for this projecct:

1. lambda = 0.2 for all of the simulations. 
2. Investigate the distribution of averages of 40 exponentials
3. Need to do a thousand simulations.


# Simulations

Set the conditions
```{r, cache=TRUE}
lambda <- 0.2
n <- 40
sims <- 1000
```

Set seed for the reproducibility, and simulate exponential distribution
```{r, cache=TRUE}
set.seed(1)
sim_exp <- replicate(sims, rexp(n, lambda))
dim(sim_exp)
```
Calculate the mean
```{r, cache=TRUE}
mean_exp <- apply(sim_exp, 2, mean)
```


#1. Sample Mean versus Theoretical Mean

Plot histogram of the means of the exponential distribution.
```{r, cache=TRUE}
hist(mean_exp, main = "Distribution of sample mean", xlab = "mean", xlim = c(2, 9), breaks = 30, col = "grey")
abline(v = mean(mean_exp), lwd = "3", col = "blue")
```


The sample mean is as follow:
```{r, cache=TRUE}
mean(mean_exp)
```

The theoretical mean of exponential distribution is 1/lambda. 

```{r, cache=TRUE}
theo_mean <- 1/lambda
theo_mean
```



#2. Sample Variance versus Theoretical Variance

The sample variance is as follow:
```{r, cache=TRUE}
var(mean_exp)
```

The theoretical variance of exponential distribution is ((1/lambda)/sqrt(n))^2.
```{r, cache=TRUE}
theo_var <- ((1/lambda)/sqrt(n))^2
theo_var
```


# Distribution

The exponential distribution of 1000 simulations is approximately normal. Due to the Central Limit Theorem, the means of the sample simulations should follow a normal distribution.

```{r, cache=TRUE}
hist(mean_exp, main = "Distribution of sample mean", xlab = "mean", xlim = c(2, 9), breaks = 30, col = "grey", probability = TRUE)

# add density plot
lines(density(mean_exp), lwd=3, col="red")  

# add normal distribution line
x <- seq(min(mean_exp), max(mean_exp), length=2*n)
y <- dnorm(x, mean=1/lambda, sd=sqrt(((1/lambda)/sqrt(n))^2))
lines(x, y, col="black", lwd=3, lty = 2)
```