---
title: "Course6-2"
author: "Sangwon Han"
date: "2018-12-18"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.


#1. Load the ToothGrowth data and explore data analyses

```{r, cache=TRUE}
data(ToothGrowth)
str(ToothGrowth)
```

```{r, cache=TRUE}
head(ToothGrowth)
```

```{r, cache=TRUE}
unique(ToothGrowth$dose)
```

#2. Provide a basic summary of the data.
```{r, cache=TRUE}
summary(ToothGrowth)
```

Plot box and whisker diagram
```{r, cache=TRUE}
par(mfrow = c(1,2))
plot(len ~ as.factor(dose), data = ToothGrowth, xlab = "Dose (mg/day)", ylab = "Tooth length")
plot(len ~ supp, data = ToothGrowth, xlab = "Supplement", ylab = "Tooth length")
```


Plot box and whisker diagram according to supplement type subgroups
```{r, cache=TRUE}
par(mfrow = c(1,2))
plot(len[ToothGrowth$supp == "OJ"] ~ as.factor(dose[ToothGrowth$supp == "OJ"]), data = ToothGrowth, xlab = "Dose (mg/day)", ylab = "Tooth length", main = "OJ", ylim = c(5,35))
plot(len[ToothGrowth$supp == "VC"] ~ as.factor(dose[ToothGrowth$supp == "OJ"]), data = ToothGrowth, xlab = "Dose (mg/day)", ylab = "Tooth length", main = "VC")
```


Plot box and whisker diagram according to dose subgroups
```{r, cache=TRUE}
par(mfrow = c(1,3))
plot(len[ToothGrowth$dose == 0.5] ~ supp[ToothGrowth$dose == 0.5], data = ToothGrowth, xlab = "Supplement", ylab = "Tooth length", main = "0.5 mg/day", ylim = c(5, 35))
plot(len[ToothGrowth$dose == 1.0] ~ supp[ToothGrowth$dose == 1.0], data = ToothGrowth, xlab = "Supplement", ylab = "Tooth length", main = "1.0 mg/day", ylim = c(5, 35))
plot(len[ToothGrowth$dose == 2.0] ~ supp[ToothGrowth$dose == 2.0], data = ToothGrowth, xlab = "Supplement", ylab = "Tooth length", main = "2.0 mg/day", ylim = c(5, 35))
```

#3. Hypothesis tests to compare tooth growth by supp and dose.

##3-1. Supplement

See whether the test meets equal variance assumption
```{r, cache=TRUE}
var.test(len~supp, data = ToothGrowth)
```
No significant difference in variance are shown.

Perform t.test function to compare two supplement groups
```{r, cache=TRUE}
t.test(len~supp, data = ToothGrowth, paired = FALSE, var.equal = TRUE)
```



##3-2. Dose

See whether the test meets equal variance assumption
```{r, cache=TRUE}
bartlett.test(len~as.factor(dose), data = ToothGrowth)
```
No significant difference in variance are shown.


Perform aov function to compare three dose groups
```{r, cache=TRUE}
ANOVA <- aov(len~as.factor(dose), data = ToothGrowth)
summary(ANOVA)
```

Then, which group shows significant difference in tooth length?
```{r, cache=TRUE}
TukeyHSD(ANOVA)
```

#4. conclusions

As shown above, we can conclude 

1. Tooth length does not differ according to supplement delivery type.
2. Tooth length does differ among the three dose levels of vitamin C.

